\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-15 paper/main.tex
final paper version, new .bst

% Why? (Interesting point #1. However, we do not know aspect #2. 
% How? (We therefore ...)
% What ...

@@

When combining cilia motion and cardiac pulsations, we recognize flow features
persisting throughout the cardiac cycle, now with vortex structures in
the anterior and middle ventricles and directional flow in the
posterior ventricle (\Crefpanels{fig:fig1}{h--j},
\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video 1}).
These characteristic patterns suggest that the flow in
the anterior and middle ventricles is primarily driven by the motile
cilia, while the cardiac pressure pulsations dominate within the
posterior ventricle and in the ducts connecting the ventricular
compartments. The CSF flow speed peaks at the time of peak caudal
cardiac pulsatile flow, when both the tangential traction induced by
the cilia and the pulsatile pressure-driven flow are aligned with the
rostrocaudal axis, reaching up to 27.9 \textmu m/s dorsally in the
posterior part of the middle ventricle. The flow Reynolds number is
$\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes
flow. In summary, the computational CSF flow model reproduces the flow
features observed in vivo~\cite{Olstad2019CiliaryDevelopment}.

@@

Interestingly, the distribution patterns differ between the
solutes. For the smallest diffusion coefficient, representing
extracellular vesicles, the advective transport by the CSF dominates diffusion: the
CSF flow structures carry over into the concentration field,
indicating that transport along the streamlines is more rapid than the
diffusion across the streamlines (\Crefpanels{fig:fig2}{b}). The
balance between diffusion and advection shifts with increasing
diffusion coefficient as the solute spreads more uniformly throughout
the ventricular geometry (\Crefpanels{fig:fig2}{b--d}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to distal regions. Computing Péclet numbers, we find that
$\mathrm{Pe}_1=664$, $\mathrm{Pe}_2=25$ and $\mathrm{Pe}_3=13$ for
$D_1$, $D_2$ and $D_3$, respectively. With all $\mathrm{Pe} > 1$, this
indicates that advection is the dominant transport mechanism on a
global scale.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein (\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video 2}).
To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F(t)
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

Turning to transport of Dendra2, removal of the
dorsal cilia again has the most pronounced effect. All mean concentration
profiles (except ROI 1, where we impose the photoconversion curve)
change noticeably (\Crefpanels{fig:fig5}{c}). Compared to the baseline
model results, we observe less solute distribution in the middle and
anterior ventricles (ROIs 2--5), and more rapid transport to the
posterior ventricle (ROI 6). On the other hand, removal of anterior
cilia does not alter the mean concentrations
(\Crefpanels{fig:fig5}{c}). Finally, when removing ventral cilia in the
middle ventricle more solute spreads towards the anterior ventricle
(ROI 5), and less solute spreads towards the posterior ventricle (ROI 6).
\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate cilia regions for which cilia paralysis is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} The total tangential force applied to the boundary with the tangential traction
    boundary condition in each of the three separate ciliated regions, as well as the total force
    (the sum of the three regions), calculated with \Cref{eq:tangential_force_calculation}. 
    \textbf{c.} Mean concentration $\cbar(t)$ for the six regions of interest (ROIs) in the three 
    paralysis scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:fig5}
\end{figure}

@@

The numerical solution schemes for the Stokes equations and the
advection-diffusion equation were implemented and solved numerically with
DOLFINx with petsc4py as linear algebra backend
\cite{dolfinx2023preprint, Dalcin2011ParallelPython}.
The simulation code is openly
available~\cite{Herlyng2025Zenodo}.
A direct solver using MUMPS~\cite{Amestoy2011Mumps} is used to solve the
linear systems resulting from the discretized Stokes equations.
For the discretized advection-diffusion equation, we use a flexible GMRES solver
with block Jacobi preconditioning \cite{fgmres, Jacobi1845UeberGleichungen}. 
We discretize the time-dependent diffusion equation (without advection)
and use the assembled linear system matrix
as the preconditioner matrix for the advection-diffusion problem.
Numerical experiments verified the accuracy of the model implementation
(\Cref{sec:appB_verification}).

@@

% Data Availability Statement
\medskip
\noindent \textbf{Data availability statement} \par
\noindent The software and data used and presented in this paper are
openly available\cite{Herlyng2025Zenodo}.

@@

% Bibliography
\bibliographystyle{ieeetr}
\bibliography{ref}

>>>>

@@

When combining cilia motion and cardiac pulsations, we recognize flow features
persisting throughout the cardiac cycle, now with vortex structures in
the anterior and middle ventricles and directional flow in the
posterior ventricle (\Crefpanels{fig:fig1}{h--j},
\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video S1}).
These characteristic patterns suggest that the flow in
the anterior and middle ventricles is primarily driven by the motile
cilia, while the cardiac pressure pulsations dominate within the
posterior ventricle and in the ducts connecting the ventricular
compartments. The CSF flow speed peaks at the time of peak caudal
cardiac pulsatile flow, when both the tangential traction induced by
the cilia and the pulsatile pressure-driven flow are aligned with the
rostrocaudal axis, reaching up to 27.9 \textmu m/s dorsally in the
posterior part of the middle ventricle. The flow Reynolds number is
$\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes
flow. In summary, the computational CSF flow model reproduces the flow
features observed in vivo~\cite{Olstad2019CiliaryDevelopment}.

@@

Interestingly, the distribution patterns differ between the
solutes. For the smallest diffusion coefficient, representing
extracellular vesicles, the advective transport by the CSF dominates diffusion: the
CSF flow structures are evident in the concentration field,
indicating that transport along the streamlines is more rapid than the
diffusion across the streamlines (\Crefpanels{fig:fig2}{b}). The
balance between diffusion and advection shifts with increasing
diffusion coefficient as solute spreads more uniformly throughout
the ventricular geometry (\Crefpanels{fig:fig2}{b--d}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to distal regions. Computing Péclet numbers, we find 
$\mathrm{Pe}_1=664$, $\mathrm{Pe}_2=25$ and $\mathrm{Pe}_3=13$ for
$D_1$, $D_2$ and $D_3$, respectively. With all $\mathrm{Pe} > 1$, this
indicates that advection is the dominant transport mechanism on a
global scale.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein (\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video S2}).
To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F(t)
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

Turning to transport of Dendra2, removal of the
dorsal cilia again has the most pronounced effect. All mean concentration
profiles (except ROI 1, where we impose the photoconversion curve)
change noticeably (\Crefpanels{fig:fig5}{c}). Compared to the baseline
model results, we observe less solute distribution in the middle and
anterior ventricles (ROIs 2--5), and more rapid transport to the
posterior ventricle (ROI 6). On the other hand, removal of anterior
cilia does not alter the mean concentrations
(\Crefpanels{fig:fig5}{c}). Finally, when removing ventral cilia in the
middle ventricle more solute spreads towards the anterior ventricle
(ROI 5), and less solute spreads towards the posterior ventricle (ROI 6).
\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate cilia regions for which cilia paralysis is simulated
    by removing the cilia tangential traction boundary condition in the respective region.
    \textbf{b.} The total force applied to the boundary with the cilia traction
    in each of the three separate ciliated regions, and the total force
    (the sum of the three regions), calculated with \Cref{eq:tangential_force_calculation}. 
    \textbf{c.} Mean concentration $\cbar(t)$ for the six regions of interest (ROIs) in the three 
    paralysis scenarios. The legend refers to the region where cilia are removed.}
    \label{fig:fig5}
\end{figure}

@@

The numerical methods for the Stokes equations and the
advection-diffusion equation were implemented and solved
with DOLFINx\cite{dolfinx2023preprint}.
The simulation code is openly
available~\cite{Herlyng2025Zenodo}.
A direct solver using MUMPS~\cite{Amestoy2011Mumps} is used to solve the
linear systems resulting from the discretized Stokes equations.
For the discretized advection-diffusion equation, we use a flexible GMRES solver
with block Jacobi preconditioning \cite{fgmres, Jacobi1845UeberGleichungen}. 
We discretize the time-dependent diffusion equation (without advection)
and use the assembled linear system matrix
as the preconditioner matrix for the advection-diffusion problem.
Numerical experiments verified the accuracy of the model implementation
(\Cref{sec:appB_verification}).

@@

% Data Availability Statement
\medskip
\noindent \textbf{Data availability statement} \par
\noindent The software and data used and presented in this paper,
as well as the supplementary videos, are
archived on Zenodo and openly available at
\href{https://doi.org/10.5281/zenodo.15194757}{https://doi.org/10.5281/zenodo.15194757}\cite{Herlyng2025Zenodo}.

@@

% Bibliography
\bibliographystyle{unsrturl}
\bibliography{ref}
